Supplementary Materials

### 2023-06-13 manuscript/supplementary.tex
supp stuff
ms changes

This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
be around 74\% for Yoruba, and 80-81\% for European and Han Chinese models
(Figure \ref{suppfig:r2_upper} This rough approximation assumes constant
demography, which for bottlenecked out-of-Africa populations assumes that the
variance in coalescence times is determined entirely by reducing the effective
population size. However, the variance in coalescence times is \emph{higher} in
bottlenecked populations (CEU and CHB) compared to Yoruba, which we confirm
with simple simulations using the \texttt{OutOfAfrica\_3G09} model from
\texttt{stdpopsim} \parencite{Gutenkunst2009-pg,Adrion2020-cf}. This higher
than expected variance would inflate $SS_\text{res}$, which would decrease

A more accurate approximate of the upper $R^2$ bound could rely on simulations
with more realistic demography.

\begin{figure}[htbp] \centering
    \includegraphics[width=0.5\textwidth]{figures/supplementary/suppfigure_R2_upper.pdf}
    \caption{A rough approximation to the theoretic $R^2$ upper bound assuming constant demography.}
  \label{suppfig:r2_upper}
\end{figure}

>>>>

This back-of-the-envelope calculation finds that the upper bound of $R^2$ would
be around 67\% for Yoruba, and 64\% for European and Han Chinese models (Figure
\ref{suppfig:drift_r2} This rough approximation assumes constant demography,
which for bottlenecked out-of-Africa populations assumes that the variance in
coalescence times is determined entirely by reducing the effective population
size. However, the variance in coalescence times is \emph{higher} in
bottlenecked populations (CEU and CHB) compared to Yoruba, which we confirm
with simple simulations using the \texttt{OutOfAfrica\_3G09} model from
\texttt{stdpopsim} \parencite{Gutenkunst2009-pg,Adrion2020-cf}. This higher
than expected variance would inflate $SS_\text{res}$, which would decrease
$R_\text{drift}^2$, bringing it closer to the observed $R_\text{LOO}^2$. A more
accurate approximation of the $R_\text{drift}^2$ could be found with
simulations with more realistic demography and varying coalescence rates along
the genome.

\begin{figure}[htbp] \centering
    \includegraphics[width=\textwidth]{figures/supplementary/suppfigure_drift_R2.pdf}
    \caption{(A) the theoretic $R_\text{drift}^2$ using different plugin
        estimators for $B_i$ under different models. This shows that among the
        best-fitting models, the predicted $R_\text{drift}^2$ varies little;
        most variation is among populations due to differing $\pi_0$. (B) The 
        $R_\text{drift}^2$ across different populations, for varying ratio of recombination
        to mutation rates (x-axis).}
  \label{suppfig:drift_r2}
\end{figure}
